D3064R0: How to Avoid OOTA Without Really Trying

### 2024-03-19 ootavacuous.tex
"Interthread Communications": Call out perftemporal.sh
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
Use "exists" consistently
Most places put "exists" in the code font, but a couple enclose it in
quotes instead.  Make them consistent with the others.

Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Finally, the reader might desire hard evidence that coe and fre
links really can go backwards in time.
We provide this evidence on x86 to demonstrate that these effects
are not confined to weakly ordered architectures.
The machine is a dual-socket system with
Intel(R) Xeon(R) Gold 6138 CPUs @ 2.00~GHz,
each socket with 20~cores and each core having a pair of hardware
threads, for a grand total of 80 hardware threads.
The code generating this data may be found in the \path{CodeSamples/cpu}
directory of the git archive of the book
``Is Parallel Programming Hard, And, If So, What Can You Do About
It?''~\cite{McKenney2018ParallelProgramming-2018-12-08a}.\footnote{
	\co{git://git.kernel.org/pub/scm/linux/kernel/git/paulmck/perfbook.git}}

@@

If a compiler makes this observation and optimizes the program by
storing a constant 17 to \co{z} rather than going through the
computations on lines~15--17 of
Listing~\ref{lst:Why rfe Instead of Tried-And-True rf?},
it will generate an executable which could produce the all-17's result
described by the ``exists'' clause.
(See Listing~\ref{lst:Why rfe Instead of Tried-And-True rf When z=17?},
where the store to \co{z} has been moved up to line~9.\footnote{
	The load from \co{y} on line~15 is retained even though it is
	not used, because \co{r3} appears in the ``exists'' clause and
	hence its final value is considered observable behavior.})
This would not be an OOTA cycle, because in either form of the program
the store to~\co{z} in \co{P0} is not semantically dependent on the
load from~\co{x}, so there is no cycle in (sdep $\cup$ rfe).
A different compiler that does not make this optimization will
generate an executable which cannot produce the all-17's result at all,
because of hardware dependencies and consequent instruction ordering
in the machine code.

>>>>

Finally, the reader might desire hard evidence that coe and fre
links really can go backwards in time.
We provide this evidence on x86 to demonstrate that these effects
are not confined to weakly ordered architectures.
The machine is a dual-socket system with
Intel(R) Xeon(R) Gold 6138 CPUs @ 2.00~GHz,
each socket with 20~cores and each core having a pair of hardware
threads, for a grand total of 80 hardware threads.
The code generating this data may be found in the \path{CodeSamples/cpu}
directory of the git archive of the book
``Is Parallel Programming Hard, And, If So, What Can You Do About
It?''~\cite{McKenney2018ParallelProgramming-2018-12-08a}.\footnote{
	\co{git://git.kernel.org/pub/scm/linux/kernel/git/paulmck/perfbook.git}}
See especially the \path{perftemporal.sh} script.

@@

If a compiler makes this observation and optimizes the program by
storing a constant 17 to \co{z} rather than going through the
computations on lines~15--17 of
Listing~\ref{lst:Why rfe Instead of Tried-And-True rf?},
it will generate an executable which could produce the all-17's result
described by the \co{exists} clause.
(See Listing~\ref{lst:Why rfe Instead of Tried-And-True rf When z=17?},
where the store to \co{z} has been moved up to line~9.\footnote{
	The load from \co{y} on line~15 is retained even though it is
	not used, because \co{r3} appears in the \co{exists} clause and
	hence its final value is considered observable behavior.})
This would not be an OOTA cycle, because in either form of the program
the store to~\co{z} in \co{P0} is not semantically dependent on the
load from~\co{x}, so there is no cycle in (sdep $\cup$ rfe).
A different compiler that does not make this optimization will
generate an executable which cannot produce the all-17's result at all,
because of hardware dependencies and consequent instruction ordering
in the machine code.
